

# My First knitr Document
Andrey Budish

## Introduction
This is some text.
Here is a code chunk.
```{r simulation, echo=TRUE}
set.seed(1)
x <- rnorm(100)
mean(x)
```

One can hide the results(look into code):  
```{r hidden, echo=FALSE, results='hide'}
set.seed(1)
x <- rnorm(100)
mean(x)
```

# Inline Text Computations
```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```
The current time is `r time`. My favorite random number is `r rand`.

# Icorporating graphics
Let's first simulate some data.
```{r simulatedata, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd=0.5)
```
Here is the scatterplot of the data.
```{r scatterplot, fig.height=4}
par(mar = c(5, 5, 1, 1), las = 1)
plot(x, y, main = "My Simulated Data")
```

# Making Tables with xtable
```{r fitmodel}
library(datasets)
data("airquality")
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```
Here is a table of regression coefficients.
```{r showtable, results="asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")
```

# Setting Global Options
``` {r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "hide")
```