---
title: "Presentation_Practice"
author: "Andrey Budish"
date: "April 15, 2017"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## My new slide
some content

***
A slide without title

## My title
### A top level subheading

#### 2nd level subheading

## Text
plain text  
*italicized text*  
**bold text**  
`code : for (i %in% 1:10)`

## Bullets
- bullet 1
- bullet 2
1. item 1
2. item 2

## R code
```{r, comment="", echo=TRUE}
head(mtcars)
```
## No code displayed
```{r, comment="", eval=FALSE, echo=TRUE}
head(mtcars)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE, fig.align='center', fig.cap="My awesome plot"}
plot(mtcars$wt, mtcars$mpg, xlab = "Weight", ylab = "MPG")
```

